Liveness-Based Garbage Collection for Lazy Languages

### 2015-12-01 paper/paper.tex
Almost final version

We shall now present a proof of soundness of the analysis presented in
Section~\ref{sec:liveness-analysis}.   Here are  the ideas  behind the
proof.
\begin{enumerate}
\item       We   augment     the        standard       semantics
  (Figure~\ref{fig:lang-semantics}) to model  a liveness-based garbage
  collection  just  before  each  $\rightsquigarrow$  transition.  The
  simulated garbage collection starts from  the root-set and inserts a
  special value  $\bot$ at each location  in the heap that  contains a
  reference that is reachable but not live. Any attempt to dereference
  such  locations during  the  transition will  result  in entering  a
  special state denoted \bang. We call the semantics after
  augmentation,  \emph{minefield semantics}.
\item \label{inline1} Assuming that a  program enters the \bang\ state,
  we  show  how to  construct,  through  inline expansion,  a  program
  without function calls which has the same minefield behavior.
\item  The final  step   shows  that  no program  without
  function calls can enter a \bang\ state. As a consequence no program
  (with or without function calls) can enter the \bang\ state. 
\end{enumerate}

@@

Simplifiers~\cite{ONeill}, described  as lightweight  daemons, improve
the efficiency of the program in general and the garbage collectors in
particular.  \warning{Most  of the  simplifications mentioned  in this
  work are subsumed by a liveness  based collector.  We feel this work
  is orthogonal  and can augment  our liveness based  collector.}  The
approach that is closest to the  method described in this paper is the
liveness-based garbage collector implemented in~\cite{karkare07liveness,asati14lgc}.  We
have extend this to handle lazy evaluation and closures.

>>>>

We shall now present a proof of soundness of the analysis presented in
Section~\ref{sec:liveness-analysis}.   Here are  the ideas  behind the
proof.
\begin{enumerate}
\item       We   augment     the        standard       semantics
  (Figure~\ref{fig:lang-semantics}) to model  a dynamic liveness analysis 
that records liveness values at each reference created during execution.
Any attempt to dereference a location with a null-liveness  
  during  a  transition will  result  in entering  a
  special state denoted \bang. We call the semantics after
  augmentation,  \emph{minefield semantics}.
\item \label{inline1} Assuming that a  program enters the \bang\ state,
  we  show  how to  construct,  through  inline expansion,  a  program
  without function calls which has the same minefield behavior.
\item  The final  step   shows  that  no program  without
  function calls can enter a \bang\ state. As a consequence no program
  (with or without function calls) can enter the \bang\ state. 
\end{enumerate}

@@

Simplifiers~\cite{ONeill}, described  as lightweight  daemons, improve
the efficiency of the program in general and garbage collectors in
particular.  \warning{Most  of the  simplifications mentioned  in this
  work are subsumed by a liveness  based collector.  We feel this work
  is orthogonal  and can augment  our liveness based  collector.}  The
approach that is closest to the  method described in this paper is the
liveness-based garbage collector implemented in~\cite{karkare07liveness,asati14lgc}.  We
have extend this to handle lazy evaluation and closures.
